\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-08-21 paper/ResponseToReviewers.tex

who told had that the burglar/loss awful was\\
who about the awful burglar/loss told had\\

These are completely different structures with even different word orders. As a result, any difference in RTs observed at the critical region might ensue from a number of different reasons that might have nothing to do with burglar/loss being or not the subject of the embedded clause.

3. The semantic manipulation is not minimal either. In this case, not only is the number of letters different (as acknowledged by the authors themselves), but even the number of syllables isn't the same. Once more, this is not a minimal comparison.''
\end{quote}

\subsubsection*{Response to R2.1}
We believe that we have addressed this concern in our Response to E.1. We summarize our response here: We have used a well-studied design that was previously published in a couple of journals, e.g., JEP:LMC. The previously published studies used exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}. The reviewer may well be right that the syntactic manipulation in the present design may not be appropriate; however, it was previously claimed in publications that this design leads to syntactic interference. Our results contribute to the literature on syntactic interference as they serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}. The reviewer's question--which exact syntactic manipulation leads to interference--is an important concern that could be the research focus of future work, but it is outside of the scope of our investigation in the present manuscript.

### 2024-08-21 paper/ResponseToReviewers.tex

3. The semantic manipulation is not minimal either. In this case, not only is the number of letters different (as acknowledged by the authors themselves), but even the number of syllables isn't the same. Once more, this is not a minimal comparison.''
\end{quote}

\subsubsection*{Response to R2.1}
We believe that we have addressed this concern in our Response to E.1. We summarize our response here: We have used a well-studied design that was previously published in a couple of journals, e.g., JEP:LMC. The previously published studies used exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}. The reviewer may well be right that the syntactic manipulation in the present design may not be appropriate; however, it was previously claimed in publications that this design leads to syntactic interference. Our results contribute to the literature on syntactic interference as they serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}. The reviewer's question--which exact syntactic manipulation leads to interference--is an important concern that could be the research focus of future work, but it is outside of the scope of our investigation in the present manuscript.

\subsection*{Comment R2.2}
\begin{quote}
``4. The difference in RTs starts at the distractor and continues through the critical verb region (in the high syntactic interference condition at least). Therefore, we cannot exclude the possibility that the effect observed at the critical verb is the same effect that originates at the distractor that is lingering through the sentence, rather than being a different effect altogether. Consequently, \textbf{I don't think that the effect at the critical region can be interpreted.}''
\end{quote}

\subsubsection*{Response to R2.2}
We agree with the reviewer. We now refrain from interpreting the results at the critical region. To emphasize this point, we added the following to the manuscript on page \pageref{only_precritical}:\\

@@

\Paste{figwhole_sentence}

\newpage

\subsection*{Comment R2.4}
\begin{quote}
``5. \textbf{Another puzzling aspect is why the effect of the semantic manipulation only manifests in the high syntactic interference condition. The authors fail to address this interaction altogether.} Instead, they repeatedly emphasize finding evidence for semantic interference and against syntactic interference, without discussing the interaction. Once more, I believe these results are difficult to interpret, but it is surprising to see such a limited discussion of the findings.''
\end{quote}

\subsubsection*{Response to R2.4}
We agree with the reviewer that visual inspection of Figure \ref{fig:whole_sentence} suggests an interaction, but the Bayes factor analysis, i.e., a formal model comparison, provided no evidence for the interaction. Therefore, we do not discuss this visual impression. See the following text in the manuscript on page \pageref{insig_interaction}:

@@

The design of the current studies repeats a design used in previous studies in German and other languages. The more novel aspects of the current studies are the use of the ERP measure, and the inclusion of a very large number of participants, allowing for more precise estimates of the syntactic and semantic interference effects. The quantitative analyses are characteristically thorough and careful.

I would support publication in JML. But I think that the paper would be stronger if there was fuller engagement with the range of possible reasons for the observed effects. This covers (i) mechanisms that could capture the lack of a syntactic interference effect, and (ii) reasons for differences between the current study and others where a syntactic interference effect was found.

The authors attribute the lack of a syntactic interference effect to the use of a syntactic feature like ``subject of the same clause", which has the effect that a subject of a different clause does not count as similar to the retrieval cue. This gives a special status to the correct subject via the choice of retrieval cues. \textbf{There may be other ways of giving special status to the correct subject, and thereby capturing the lack of syntactic interference. For example, if the subject of the current clause is encoded in a special status, e.g., with boosted activation, or with its needs prioritized, then this could have a similar effect.}''
\end{quote}

\subsubsection*{Response to R3.1}
The reviewer's proposal could well be right, that the subject of the current clause receives an activation boost. However, the mechanism for implementing this idea computationally might not be trivial. Let's assume that the subject is encoded with boosted activation compared to for example an object. As the sentence progresses the subject's activation decays (we could even assume that it decays more slowly than other memory representations due to its special status) and other nouns are processed / encoded. At one point within a long sentence, the subject's activation will be lower than the activation of other nouns in memory. If we assume that a long sentence has multiple embedded clauses, we would need to assume that either the matrix sentence subject activation is downgraded when an embedded sentence is entered or that the embedded sentence subject gets an even higher activation boost than the matrix sentence subject. Then when the embedded sentence is finished and we return back to the matrix sentence, the matrix sentence subject would need to receive another activation boost to return to its special status. Essentially, this proposal condenses down to a clause tracking mechanism with a modulation of activation levels. In our opinion, it is more parsimonious to implement the clause tracking in the memory encoding and retrieval operations than the activation levels.

### 2024-08-21 paper/ResponseToReviewers.tex

\begin{quote}
    \Paste{semantic_features_encoding}    
\end{quote}

\subsection*{Comment R3.4}
\begin{quote}
``\textbf{The authors offer only limited discussion of why their results differed from those of other studies. To some degree, the paper seems to take the position that since the current studies had huge numbers of participants and hence high statistical power, this study reveals the true state of the world, and other studies should be taken less seriously because their results were less reliable.}''
\end{quote}

\subsubsection*{Response to R3.4}
As described in our Response to R1.4, we have added a detailed comparison of our results to the previous results. The main difference is that our data showed no evidence for syntactic interference in the pre-critical region, while the previous studies did show support for it. We have created a new subsection in our General Discussion providing an extended discussion of the potential reasons for the differences between our and previous results. This discussion did not lead to a definite answer. We provide the relevant new text passages in our responses below (Responses to R3.7 and R3.8).

We think that we have addressed the concerns regarding high- vs.\ lower-powered studies and contradictory results between studies in our Response to R1.5. We do not claim that our results are correct and previous claims have been false. We feel that progress in sentence processing research will be greatly facilitated by carrying out adequately powered studies in the future that can
distinguish between theoretical claims. Our work aims to be a contribution in this direction. As discussed earlier in this response letter, the contradiction between published results even when using the same design (and even the same items) is not an anomaly in sentence processing research.

### 2024-08-21 paper/main.tex
fixed formatting
Resolved Shravan's comments on modeling section

\textcolor{blue}{The procedure for computing Bayes factors is as follows. The Bayes factor computation required us to estimate the marginal likelihood of the competing models. We used the Monte Carlo integration method to estimate the marginal likelihood of each model. First, we chose an importance density for the parameter $F$ from which we draw $0.2$ million samples of $F$. Then, for each of these proposed parameter values, we compute the likelihood and prior density. The estimated marginal likelihood of a model will be given by}

\begin{equation}
ML = \frac{1}{n} \sum_{i=1}{n} \mathcal{L}(F_i|y) p(F_i)  \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
\end{equation}

\noindent \textcolor{blue}{where $y$ is the observed data, i.e., the observed semantic interference effect in our case, and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $. 
The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$ and $p(F_i)$ gives the prior density of $F_i$. We compute marginal likelihoods under different prior assumptions $p(F)$.}

\textcolor{blue}{The Lewis and Vasishth (2005) model is a complex, non-deterministic model which can be simplified such that a likelihood function represents the model's generative process  \parencite[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}. However, it is also possible to carry out model comparison without recasting the model as a likelihood function. In this case, model comparison is carried out using Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood}. Under this approach, the model-generated data conditional on $F=F_i$ is compared with the actually observed data $y$. If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa. The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weighs a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.}

>>>>

\textcolor{blue}{The procedure for computing Bayes factors is as follows. The Bayes factor computation required us to estimate the marginal likelihood of the competing models. We used the Monte Carlo integration method to estimate the marginal likelihood of each model. For Monte Carlo integration, we use the importance sampling estimator where a large number of number proposals are obtained from an importance density $q(F)$ and the average $likelihood \times prior$ is calculated across all these proposals to obtain the approximate marginal likelihood. The importance density should be chosen such that it covers the entire parameter space where $likelihood \times prior$ is non-zero and does not draw too many samples from the space where $likelihood \times prior$ is zero. We estimated the marginal likelihoods using Monte Carlo importance sampling as follows.

First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we use a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density. This importance density allows us to obtain more proposals for smaller values of $F$ (for F<0.03) while maintaining fatter tails for the larger values of $F$ ($F>0.03$); effectively, the whole parameter space of $F$ between $0$ and $0.05$ is explored to draw $0.2$ million proposals. Then, for each of these proposed parameter values, we compute the likelihood and prior density. The estimated marginal likelihood of a model will be given by}

\begin{equation}
ML = \frac{1}{n} \sum_{i=1}{n} \frac{\mathcal{L}(F_i|y) p(F_i)}{q(F_i)} \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
\end{equation}

\noindent \textcolor{blue}{where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case, and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $. 
The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$. We compute marginal likelihoods under different prior assumptions $p(F)$.}

\textcolor{blue}{The Lewis and Vasishth (2005) model is a complex, non-deterministic model which can be simplified such that a likelihood function represents the model's generative process  \parencite[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}. However, it is also possible to carry out model comparison without recasting the model as a likelihood function. In this case, model comparison is carried out using Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood}. Under this approach, the model-generated data conditional on $F=F_i$ is compared with the actually observed data $y$. If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa. The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weighs a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$. We choose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$. The $\delta$ parameter determines how good is the approximation: the lower is the value of the $delta$, the better is the approximation of the likelihood. However, if $\delta$ is too low, the computational demand increases because too many proposals would end up with zero likelihood. We choose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising the computational efficiency.}
